---
title: "group_2_dplyr"
author: "Sijun Zhang"
date: "2019/12/4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R using Dplyr and glmnet

## Perparing Data

First we will import all 2015-2016 NHANES and check if the data is perpared in the common data folder. If not, we will download the data from the website.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(Hmisc)
library(SASxport)
```


```{r}
# read in the  data: -----------------------------------------------------------
## This data will be used in the question. 
url_base <- 'https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/'

demo_file <- '../../DATA/DEMO_I.XPT'
if ( !file.exists(demo_file) ) {
  demo_url <- sprintf('%s/DEMO_I.XPT', url_base)
  demo <- sasxport.get(demo_url)
  write.xport(demo, file = demo_file)
} else {
  demo <- sasxport.get(demo_file)
}

dr1_file <- '../../DATA/DR1TOT_I.XPT'
if ( !file.exists(dr1_file) ) {
  dr1_url <- sprintf('%s/DR1TOT_I.XPT', url_base)
  dr1 <- sasxport.get(dr1_url)
  write.xport(dr1, file = dr1_file)
} else {
  dr1 <- sasxport.get(dr1_file)
}

dr2_file <- '../../DATA/DR2TOT_I.XPT'
if ( !file.exists(demo_file) ) {
  dr2_url <- sprintf('%s/DR2TOT_I.XPT', url_base)
  dr2 <- sasxport.get(dr2_url)
  write.xport(dr2, file = dr2_file)
} else {
  dr2 <- sasxport.get(dr2_file)
}

bpx_file <- '../../DATA/BPX_I.XPT'
if ( !file.exists(bpx_file) ) {
  bpx_url <- sprintf('%s/BPX_I.XPT', url_base)
  bpx <- sasxport.get(bpx_url)
  write.xport(bpx, file = bpx_file)
} else {
  bpx <- sasxport.get(bpx_file)
}
```

Then we select only the responents with both two day data collected with no missing value and then take average of them as the average nutrition intake.

Meanwhile, we also standarized the predictors and then merge the data together. We take the systolic and diastolic blood pressures as the responses that are needed to be averaged for each respondent. 

```{r}
dr1 = dr1 %>%
  transmute(seqn = seqn, alco = as.numeric(dr1talco),
            water = as.numeric(dr1.320z), caff = as.numeric(dr1tcaff),
            sodi = as.numeric(dr1tsodi), fat = as.numeric(dr1ttfat), 
            sugr = as.numeric(dr1tsugr), iron = as.numeric(dr1tiron),
            fibe = as.numeric(dr1tfibe), prot = as.numeric(dr1tprot) ) %>%
  gather(key = "meas", value = "day1", -seqn )

dr2 = dr2 %>%
  transmute(seqn = seqn, alco = as.numeric(dr2talco),
            water = as.numeric(dr2.320z), caff = as.numeric(dr2tcaff), 
            sodi = as.numeric(dr2tsodi), fat = as.numeric(dr2ttfat), 
            sugr = as.numeric(dr2tsugr), iron = as.numeric(dr2tiron),
            fibe = as.numeric(dr2tfibe), prot = as.numeric(dr2tprot) ) %>%
  gather(key = "meas", value = "day2", -seqn )

dr = dr1 %>%
  left_join(dr2, by = c('seqn', 'meas'))  %>%
  gather(key = "svy_day", value = "value", day1:day2) %>%
  spread(key = "meas", value = "value" ) %>%
  group_by(seqn) %>%
  summarise(alco = mean(alco), caff = mean(caff),
            fat = mean(fat), fibe = mean(fibe),
            iron = mean(iron), prot = mean(prot),
            sodi = mean(sodi), sugr = mean(sugr), water = mean(water))

demo = demo %>%
  # transmute(seqn = seqn, age = ridageyr, pir = indfmpir, gender = riagendr)
  transmute(seqn = seqn, gender = riagendr)

bpx = bpx %>%
  # most people havn't test 4 so we just ignore it
  transmute(seqn = seqn, bpxsy_avg = (bpxsy1+bpxsy2+bpxsy3)/3,
            bpxdiff_avg = ((bpxsy1-bpxdi1)+(bpxsy2-bpxdi2)+(bpxsy3-bpxdi3))/3)

bpx_dr = dr %>%
  left_join(bpx, by="seqn")


df = bpx_dr %>%
  left_join(demo, by = "seqn") %>%
  drop_na()

# standardizing process
df_scale = scale(df[,2:10])

df = df %>%
  select(seqn, bpxsy_avg, bpxdiff_avg, gender) %>%
  transmute(seqn, bpxsy_avg, bpxdiff_avg, gender = gender-1.5) %>%
  cbind(df_scale) %>%
  mutate(gender_alco = gender*alco,
         gender_caff = gender*caff,
         gender_fat = gender*fat,
         gender_fibe = gender*fibe,
         gender_iron = gender*iron,
         gender_prot = gender*prot,
         gender_sodi = gender*sodi,
         gender_sugr = gender*sugr,
         gender_water = water*gender)
```

Then, in order to compare the different data analysis tools, the lasso analysis will be done in the Python

```{r}
write.csv(df, file = "group_2_df.csv", row.names = FALSE)
```

## LASSO in Python
```{python eval=FALSE}
import numpy as np
import sklearn
import pandas as pd
import glmnet_py
from glmnet import glmnet
import scipy
```


